---
title: "run_analysis"
---

Checks for the dataset and if it does not exist, downloads and extracts the files.
```{r}
dir <- "UCI HAR Dataset"
if (!file.exists(dir)) {
    temp <- tempdir()
	url <- "http://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip"
	file <- basename(url)
	download.file(url, file)
	unzip(file)
	unlink(temp)
}
```
Filepaths of the files are set.
```{r}
testSet <- paste(dir, "/test/X_test.txt", sep = "")
testLabel <- paste(dir, "/test/y_test.txt", sep = "")
testSubject <- paste(dir, "/test/subject_test.txt", sep = "")
trainSet <- paste(dir, "/train/X_train.txt", sep = "")
trainLabel <- paste(dir, "/train/y_train.txt", sep = "")
trainSubject <- paste(dir, "/train/subject_train.txt", sep = "")
features <- paste(dir, "/features.txt", sep = "")
activityLabel <- paste(dir, "/activity_labels.txt", sep = "")
```
Files are read to separate dataframes.
```{r}
testSet <- read.table(testSet, header=F)
testLabel <- read.table(testLabel, header=F)
testSubject <- read.table(testSubject, header=F)
trainSet <- read.table(trainSet, header=F)
trainLabel <- read.table(trainLabel, header=F)
trainSubject <- read.table(trainSubject, header=F)
features <- read.table(features, header=F)
activityLabel <- read.table(activityLabel, header=F)
```
The test dataset is prepared with appropriate descriptive variable names and activity names, 
all of which are generated from the label, subject, features and activityLabel Files.
```{r}
testLabelDesc <- as.character(testLabel[1])
for (i in 1:nrow(testLabel)) {
	testLabelDesc[i] <- as.character(activityLabel[testLabel[i, 1], 2])
}
colnames(testSet) <- features[[2]]
testSet <- cbind(testSubject[1], testLabelDesc, testSet)
colnames(testSet)[1] <- "subjectID"
colnames(testSet)[2] <- "activity"
```
The train dataset is prepared with appropriate descriptive variable names and activity names, 
all of which are generated from the label, subject, features and activityLabel Files.
```{r}
trainLabelDesc <- as.character(trainLabel[1])
for (i in 1:nrow(trainLabel)) {
	trainLabelDesc[i] <- as.character(activityLabel[trainLabel[i, 1], 2])
}
colnames(trainSet) <- features[[2]]
trainSet <- cbind(trainSubject[1], trainLabelDesc, trainSet)
colnames(trainSet)[1] <- "subjectID"
colnames(trainSet)[2] <- "activity"
```
Merges the training and the test sets to create one data set.
```{r}
dataSet <- rbind(trainSet, testSet)
```
Extracts only the measurements on the mean and standard deviation for each measurement. 
```{r}
index = grep(".*mean\\(\\)|.*std\\(\\)", names(dataSet))
dataSet2 <- dataSet[ , index]
dataSet2 <- cbind(dataSet[1:2], dataSet2)
write.table(dataSet2, "dataSet2.txt")
```
From the data set in step 4, creates a second, independent tidy data set with the 
average of each variable for each activity and each subject.
```{r}
library(data.table)
dataSet3 <- data.table(dataSet2)
dataSet3 <- dataSet3[,lapply(.SD, mean),by=c("subjectID", "activity")]
write.table(dataSet3, "dataSet3.txt")
```
